A Temporal Logic Based Framework for Intrusion Detection

### 2023-09-04 Paper/main.tex
update

According to~\cite{barringer2004rule}, the definitions of $evaluate$, $update$ and $value$ on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
%evaluate
\begin{align*}
    evaluate(\underline{\text{true}},\alpha_i) & =\ \underline{\text{true}} \\
    evaluate(\underline{\text{false}},\alpha_i) & =\ \underline{\text{false}} \\ 
    evaluate(exp,\alpha_i) & =\ \text{value of } exp\text{ in } \alpha_i \\
    evaluate(F_1\ op\ F_2,\alpha_i) & =\  evaluate(F_1,\alpha_i)\ op\ evaluate(F_2,\alpha_i)\\
    evaluate(\neg{F},\alpha_i) & =\ \neg{evaluate(F,\alpha_i)}  \\
    evaluate(\bigcirc F,\alpha_i) & =\ update(F,\alpha_i)  \\
    evaluate(F_1 \cdot F_2,\alpha_i) & =
    \begin{cases}
        evaluate(F_1,\alpha_i)\cdot F_2 \quad \text{if } value(F_1)=\ \underline{\text{false}} \\
        evaluate(F_1,\alpha_i)\cdot F_2 \lor evaluate(F_2,\alpha_i)\quad \text{else} 
    \end{cases}
\end{align*}

\noindent\begin{minipage}{0.3\linewidth}
    \begin{flalign*}
        \text{value}(\underline{\text{true}}) & = \underline{\text{true}} \\
        \text{value}(\underline{\text{false}}) & = \underline{\text{false}} \\ 
        \text{value}(\text{exp}) & = \underline{\text{false}} \\
        \text{value}(F_1\ \text{op}\ F_2) & = \text{value}(F_1)\ \text{op}\ \text{value}(F_2) \\
        \text{value}(\neg{F}) & = \neg{\text{value}(F)}  \\
        \text{value}(\bigcirc F) & =
        \begin{cases}
            F & \text{if at the beginning of a trace} \\
            \underline{\text{false}} & \text{if at the end of a trace} \\
        \end{cases}\\
        \text{value}(F_1 \cdot F_2,\alpha_i) & = \text{value}(F_1)\land \text{value}(F_2)
    \end{flalign*}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
    \begin{align*}
        \text{update}(\underline{\text{true}},\alpha_i) & = \underline{\text{true}} \\
        \text{update}(\underline{\text{false}},\alpha_i) & = \underline{\text{false}} \\ 
        \text{update}(\text{exp},\alpha_i) & = \text{exp}\\
        \text{update}(F_1\ \text{op}\ F_2,\alpha_i) & = \text{update}(F_1,\alpha_i)\ \text{op}\ \text{update}(F_2,\alpha_i) \\
        \text{update}(\neg{F},\alpha_i) & = \neg{\text{update}(F,\alpha_i)}  \\
        \text{update}(\bigcirc F,\alpha_i) & = \text{update}(F,\alpha_i)  \\
        \text{update}(F_1 \cdot F_2,\alpha_i) & = \text{update}(F_1,\alpha_i)\cdot F_2
    \end{align*}
\end{minipage}

>>>>

According to~\cite{barringer2004rule}, the definitions of $evaluate$, $update$ and $value$ on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
%evaluate
\begin{align*}
    evaluate(\underline{\text{true}},\alpha_i) & =\ \underline{\text{true}} \\
    evaluate(\underline{\text{false}},\alpha_i) & =\ \underline{\text{false}} \\ 
    evaluate(exp,\alpha_i) & =\ \text{value of } exp\text{ in } \alpha_i \\
    evaluate(F_1\ op\ F_2,\alpha_i) & =\  evaluate(F_1,\alpha_i)\ op\ evaluate(F_2,\alpha_i)\\
    evaluate(\neg{F},\alpha_i) & =\ \neg{evaluate(F,\alpha_i)}  \\
    evaluate(\bigcirc F,\alpha_i) & =\ update(F,\alpha_i)  \\
    evaluate(F_1 \cdot F_2,\alpha_i) & =
    \begin{cases}
        evaluate(F_1,\alpha_i)\cdot F_2 \quad \text{if } value(F_1)=\ \underline{\text{false}} \\
        evaluate(F_1,\alpha_i)\cdot F_2 \lor evaluate(F_2,\alpha_i)\quad \text{else} 
    \end{cases} \\ \\
%\end{align*}
%\noindent\begin{minipage}[l]{\textwidth}
   % \raggedright
    %\begin{align*}
        \text{value}(\underline{\text{true}}) & = \underline{\text{true}} \\
        \text{value}(\underline{\text{false}}) & = \underline{\text{false}} \\ 
        \text{value}(\text{exp}) & = \underline{\text{false}} \\
        \text{value}(F_1\ \text{op}\ F_2) & = \text{value}(F_1)\ \text{op}\ \text{value}(F_2) \\
        \text{value}(\neg{F}) & = \neg{\text{value}(F)}  \\
        \text{value}(\bigcirc F) & =
        \begin{cases}
            F & \text{if at the beginning of a trace} \\
            \underline{\text{false}} & \text{if at the end of a trace} \\
        \end{cases}\\
        \text{value}(F_1 \cdot F_2,\alpha_i) & = \text{value}(F_1)\land \text{value}(F_2) \\ \\
   % \end{align*}
%\end{minipage}
%\hfill
%\begin{minipage}[r]{\textwidth}
   % \raggedleft
    %\begin{align*}
        \text{update}(\underline{\text{true}},\alpha_i) & = \underline{\text{true}} \\
        \text{update}(\underline{\text{false}},\alpha_i) & = \underline{\text{false}} \\ 
        \text{update}(\text{exp},\alpha_i) & = \text{exp}\\
        \text{update}(F_1\ \text{op}\ F_2,\alpha_i) & = \text{update}(F_1,\alpha_i)\ \text{op}\ \text{update}(F_2,\alpha_i) \\
        \text{update}(\neg{F},\alpha_i) & = \neg{\text{update}(F,\alpha_i)}  \\
        \text{update}(\bigcirc F,\alpha_i) & = \text{update}(F,\alpha_i)  \\
        \text{update}(F_1 \cdot F_2,\alpha_i) & = \text{update}(F_1,\alpha_i)\cdot F_2
    \end{align*}
%\end{minipage}
Let us introduce an example to better understand the algorithm. Given the following rules 
\begin{align*}
    &\underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F)  = F \land \bigcirc \text{Always}(F) \\
    &\underline{\text{min}}\ \text{EvTimedLogout}(\underline{\text{string }} k, \underline{\text{double }} t, \underline{\text{double }} \delta) = (\text{time} - t \leq \delta) \\
    &\quad\quad\quad\quad \land ((action = \text{logout} \land userId = k) \lor \bigcirc \text{EvTimedLogout}(k, t, \delta)) \\
    &\underline{\text{mon}}\ M_3  = \text{Always}((action = \text{login}) \rightarrow \text{EvTimedLogout}(userId,time, 100))
\end{align*}
and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\text{"Lori"},action:\text{login},time:\text{17.0}\}$ and $\alpha_1=\{userId:\text{"Lori"},action:\text{logout},time:\text{150.0}\}$, the initial formula $M_3$ gets modified. At time $1$, when $\alpha_1$ is analyzed, $M_3$ changes as follows:
\begin{align*}
    &evaluate(M_3,\alpha_1)  = \text{EvTimedLogout("Lori",17.0, 100)} \land \\ 
    &\quad\quad\quad\quad \text{Always}((action = \text{login}) \rightarrow \text{EvTimedLogout}(userId,time, 100))
\end{align*}
At the second event the predicate $(time-t \leq \delta)$ gets instantiated to $(150.0-17.0 \leq 100)$ which is false; thus the whole formula becomes false, indicating that the two-event trace violates the property $M_3$.

It's essential to emphasize that the logic EAGLE operates with a \textit{finite} trace semantics; the monitoring algorithm, as written before, can check whetet the formula is satisfied or not only at the conclusion of a trace. However, in intrusion detection scenarios, the notion of the 'end of a trace' makes no sense since event sequences can theoretically be infinite. In such cases, the aim is to trigger an alarm as soon as a property is violated and this is done by continuously checking the formula's satisfaction status after each event.
